---
title: "How to Use Probability Distributions"
output: learnr::tutorial
runtime: shiny_prerendered
---
```{r setup, include=FALSE}
library(learnr)
library(learnrhash)
library(tidyverse)
library(gradethis)
library(here)
tutorial_options(
  # use gradethis for checking
  exercise.checker = gradethis::grade_learnr
  )
knitr::opts_chunk$set(echo = FALSE)
```

## Generate random values from a distribution

You can generate random values from a variety of distributions including Normal, t, Exponential, and Gamma. Run the following chunk a few different times.

```{r rand, exercise = TRUE}
rnorm(2, mean = 0, sd = 1)

rt(2, df = 5)
```

If you want the random generation to be reproducible, you need to set your seed.

```{r randseed, exercise = TRUE}
set.seed(11435)

rnorm(2, mean = 0, sd = 1)

rt(2, df = 5)
```

## Calculate probabilities

You can calculate probabilities under a particular probability distribution. 

```{r prob, exercise = TRUE}
pnorm(1.96, mean = 0, sd = 1, lower.tail = TRUE)
pnorm(-1.96, mean = 0, sd = 1, lower.tail = TRUE)

pt(0.25, df = 5, lower.tail = TRUE)
```


## Find quantiles

You can also calculate quantiles of particular probability distributions.

```{r quant, exercise = TRUE}
qnorm(0.975, mean = 0, sd = 1, lower.tail = TRUE)
qnorm(0.025, mean = 0, sd = 1, lower.tail = TRUE)

qt(0.6, df = 5, lower.tail = TRUE)
```


## Your Turn

Suppose $Y \sim N(\mu = 2.7, \sigma^2 = 1.3)$. Find $P(Y \geq 3.3)$.

```{r normprob, exercise = TRUE}
_norm(___, mean = ___, sd = ___, lower.tail = ___)
```

```{r normprob-solution}
pnorm(3.3, mean = 2.7, sd = sqrt(1.3), lower.tail = FALSE)
```

```{r normprob-check}
# check code
gradethis::grade_code()
```

Find the Interquartile Range

```{r normiqr, exercise = TRUE}
_norm(___, mean = ___, sd = ___, lower.tail = ___) -
_norm(___, mean = ___, sd = ___, lower.tail = ___)
```

```{r normiqr-solution}
qnorm(0.75, mean = 2.7, sd = sqrt(1.3), lower.tail = FALSE) - 
qnorm(0.25, mean = 2.7, sd = sqrt(1.3), lower.tail = FALSE)
```

```{r normiqr-check}
# check code
gradethis::grade_code()
```

Performance ratings at a certain corporation are Normally distributed with a mean of 82.5 and a standard deviation of 9.2. Policy states that the top 15% of performers will earn a 3% raise. What is the lowest performance rating that will guarantee an employee a raise?

```{r perf, exercise = TRUE}
_norm(___, mean = ___, sd = ___, lower.tail = ___)
```

```{r perf-solution}
qnorm(0.15, mean = 82.5, sd = 9.2, lower.tail = FALSE)
```

```{r perf-check}
# check code
gradethis::grade_code()
```


## Submit progress

As you have gone through this tutorial, your progress has been kept track of in something called a "hash". I can use this to make sure you have executed all of the code chunks and answered all of the questions. I will use this to give you a completion grade for this assignment. 

If a Google Form window does not appear, you can access it [here](https://docs.google.com/forms/d/e/1FAIpQLSeW9exymQeRir3-xsUfzm6JPVWB9XHaKHkAa-hpR0ne1g7gyw/viewform).

```{r context="server"}
learnrhash::encoder_logic(strip_output = T)
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(
  ui_before = div(strong("Submit your hash in the form below."), br(), br()),
  ui_after  = learnrhash::iframe_ui(
    src = "https://docs.google.com/forms/d/e/1FAIpQLSeW9exymQeRir3-xsUfzm6JPVWB9XHaKHkAa-hpR0ne1g7gyw/viewform", 
    width="900px", height= "1000px"
  )
)
```



