---
title: "How to Use Loops"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(learnrhash)
library(tidyverse)
library(gradethis)
library(here)
tutorial_options(
  # use gradethis for checking
  exercise.checker = gradethis::grade_learnr
  )
knitr::opts_chunk$set(echo = FALSE)
```


# Getting started - Loops

A simple for loop has the following structure:

```{r loopEx, exercise = TRUE}
n <- 5
for(i in 1:n){
  print(1)
}
```

This is read "for index $i$ in the sequence 1 to $n$, do the following" (whatever is in the curly braces). 

```{r loop1}
quiz(
question("What would the for loop above do",
    answer("Print 5 one time."),
    answer("Print 5 five times"),
    answer("Print 1 five times", correct = T),
    answer("Print 1 one time."),
    allow_retry = TRUE,
    random_answer_order = F
  )
)
```

## Getting started - Loops cont.

In this class, we will often want to work with probability distributions. 

```{r loopEx2, exercise = TRUE}
n <- 5
sample_size <- 3
for(i in 1:n){
  x <- rnorm(n = sample_size, mean = 0, sd = 1)
}
```

```{r loop2}
quiz(
question("What is in x at the end of the above for loop",
    answer("3 numbers drawn from the specified normal distribution.", correct = T),
    answer("15 numbers drawn from the specified normal distribution."),
    answer("Nothing, x is empty."),
    answer("1 number drawn from the specified normal distribution"),
    allow_retry = TRUE,
    random_answer_order = F
  )
)
```


## Sums

$$\sum_{i=0}^4 3i$$

```{r sum, exercise = TRUE}
total <- 0 ## start the total at zero
for(j in 0:4){ ## for index j in 0 through 4
  
  total <- total + 3*j ## overwrite total by adding 3*j 
  
  
}
total
class(total) ## what type of object is this
length(total)
dim(total) ## not relevant for this type
```


## Products

Go through the code annotation process for the following code chunk.

$$\prod_{i=1}^4 2i$$

```{r prod, exercise = TRUE}
total <- 1 ## do you understand why this starts at 1?
for(i in 1:4){
  
  total <- total * 2*i
  
}
total
class(total)
length(total)
```

## Simulate and Store Data


```{r sim, exercise = TRUE}
num_sim <- 5 
sample_size <- 3
stored_data <- matrix(NA, nrow = num_sim, ncol = sample_size)
## set up a place to store things
## set up empty matrix of NAs with 5 rows and 3 columns
for(i in 1:num_sim){ ## do something 5 times
  x <- rnorm(n = sample_size, mean = 0, sd = 1) 
  ## draw 3 random numbers from N(0,1)
  stored_data[i,] <- x
  ## over ride the ith row in stored_data with the 3 values we drew
}
stored_data ## print results
```

```{r quiz1}
quiz(
question("What does stored_data look like when it is first created?",
    answer("A matrix of random numbers with 5 rows and 3 columns."),
    answer("A matrix of random numbers with 3 rows and 5 columns."),
    answer("A matrix of NAs with 5 rows and 3 columns.", correct = TRUE),
    answer("A matrix of NAs with 3 rows and 5 columns."),allow_retry = TRUE,
    random_answer_order = T
  ),
question("What function should we use to determine the size of stored_data?",
    answer("rep"),
    answer("length"),
    answer("dim", correct = TRUE),
    answer("c()"),allow_retry = TRUE,
    random_answer_order = T
  )
)
```

## Simulate and Store Results

```{r sim2, exercise = TRUE}
num_sim <- 5
sample_size <- 3
stored_means <- rep(NA, length.out = num_sim) ## we haven't seen this before, so be sure to check the help page
for(i in 1:num_sim){
  x <- rnorm(n = sample_size, mean = 0, sd = 1)
  stored_means[i] <- mean(x)
  }
stored_means
```

```{r quiz2}
quiz(
question("What does stored_means look like when it is first created?",
    answer("A matrix of NAs with 5 rows and 3 columns."),
    answer("A vector of NAs that has 3 elements in it."),
    answer("A vector of NAs that has 5 elements in it.", correct = TRUE),
    answer("5 repeated 3 times."),allow_retry = TRUE,
    random_answer_order = T
  ),
question("What function should we use to determine how many elements are in stored_means?",
    answer("rep"),
    answer("dim"),
    answer("length", correct = TRUE),
    answer("c()"),allow_retry = TRUE,
    random_answer_order = T
  )
)
```

## Your Turn

Suppose we are interested in approximating the sampling distribution of the mean of an Exponential random variable. 

```{r getMean, exercise = TRUE, exercise.eval = FALSE}
lambda <- 1 
n <- 10 
num_rep <- 1000 ## how many times we want to repeat the simulation, something fairly large
store_mean <- ___(NA, ___)
for (i in 1:___) {
  test <- rexp(n = ___, rate = ___)
  store_mean[___] <- ___(test)
}
mean(___) ## to approximate the expected value
var(___) ## to approximate the variance 
```

```{r getMean-solution}
lambda <- 2 
n <- 10 
num_rep <- 1000 ## how many times we want to repeat the simulation, something fairly large
store_mean <- rep(NA, num_rep)
for (i in 1:num_rep) {
  test <- rexp(n = n, rate = lambda)
  store_mean[i] <- mean(test)
}
mean(store_mean) ## to approximate the expected value
var(store_mean) ## to approximate the variance
```

```{r getMean-check}
# check code
gradethis::grade_code()
```


## Submit progress

As you have gone through this tutorial, your progress has been kept track of in something called a "hash". I can use this to make sure you have executed all of the code chunks and answered all of the questions. I will use this to give you a completion grade for this assignment. 

If a Google Form window does not appear, you can access it [here](https://docs.google.com/forms/d/e/1FAIpQLSeW9exymQeRir3-xsUfzm6JPVWB9XHaKHkAa-hpR0ne1g7gyw/viewform).

```{r context="server"}
learnrhash::encoder_logic(strip_output = T)
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(
  ui_before = div(strong("Submit your hash in the form below."), br(), br()),
  ui_after  = learnrhash::iframe_ui(
    src = "https://docs.google.com/forms/d/e/1FAIpQLSeW9exymQeRir3-xsUfzm6JPVWB9XHaKHkAa-hpR0ne1g7gyw/viewform", 
    width="900px", height= "1000px"
  )
)
```



