---
title: "How to Subset Data"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(learnrhash)
library(tidyverse)
library(gradethis)
library(here)
tutorial_options(
  # use gradethis for checking
  exercise.checker = gradethis::grade_learnr
  )
knitr::opts_chunk$set(echo = FALSE)
```

## Vectors

### Overview 

The `c()` function *concatenates* items (of homogenous type) into a vector; for example, `myvector = c(1, 3, 6, 9)` creates a vector that has four elements. To learn about how many elements in the vector there are, you can use the function `length`. Read the help file for the function for more guidance.

```{r myvec, exercise = TRUE}
myvector <- c(1, 3, 6, 9)
myvector
length(myvector)
```


To access a particular element, you can use square brackets.

```{r myvecz, exercise = TRUE}
myvector <- c(1, 3, 6, 9)
myvector[2]
myvector[c(2,3)]
```

### Exercise

Use `c()` to create a vector called `myvector` that contains all the perfect squares (1, 4, 9, 16, $\dots$) less than or equal to 100.

```{r makeVec, exercise = TRUE}

```

```{r makeVec-solution}
myvector <- c(1, 4, 9, 16, 25, 36, 49, 64, 81, 100)
```

```{r makeVec-check}
# check code
gradethis::grade_code()
```

## Matrices

What are the dimensions of the following matrix? Use `dim` to find out. Read the help file for the function for more guidance on what this function returns if it isn't clear from the following example.

```{r makeMatrix, exercise = TRUE}
mymatrix <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3)
dim(___)
```

```{r makeMatrix-solution}
mymatrix <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3)
dim(mymatrix)
```

```{r makeMatrix-check}
# check code
gradethis::grade_code()
```


We can access specific rows of a matrix with `[,]` - for example, `mymatrix[2,]` will present the second row of the matrix, and `mymatrix[c(2,3),]` will yield a new matrix with the second and third rows of the original only.


```{r makeMatrix2, exercise = TRUE}
mymatrix <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3)
mymatrix[c(2,3),]
```

Assign the first row of `mymatrix` to a variable named `row1`. Use the function `summary()` to  find the minimum and maximum of this row. Consult the help files for `summary` if necessary.

```{r matrixWork, exercise = TRUE}
mymatrix <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3)
___ <- mymatrix[___]
summary(___)
```

```{r matrixWork-solution}
mymatrix <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3)
row1 <- mymatrix[1,]
summary(row1)
```

```{r matrixWork-check}
# check code
gradethis::grade_code()
```

To get a single element in a matrix, we use row and column numbers, such as `mymatrix[r,c]`. This will also extract a submatrix if multiple entries are used for each.

```{r matrixWork2, exercise = TRUE}
mymatrix <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3)
mymatrix[1,2]
```

Display the submatrix that contains the first two rows and second column of `mymatrix`. 


```{r matrixWork3, exercise = TRUE}
mymatrix <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3)
mymatrix[___, ___]
```

```{r matrixWork3-solution}
mymatrix <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3)
mymatrix[c(1,2), 2]
```

```{r matrixWork3-check}
# check code
gradethis::grade_code()
```

## Logicals


We can also ask questions about our data/variables. The `==` sign is a True/False test of equality. e.g. `row2 == 1` returns `TRUE` for every element in row2 that is equal to 1, and `FALSE` for elements different than 1. We can do similar things with the symbols `<`, `<=`, `>`, `>=`. 

```{r log, exercise = TRUE}
row2 <- c(1, 2, 3, 1)
row2 == 1
```


You can also combine two conditions with the AND (`&`) and OR (`|`) operators.

```{r log2, exercise = TRUE}
row2 <- c(1, 2, 3, 1)
col3 <- c(1, 1, 3, 2)
row2 >= 100 & col3 == 1
row2 == 1 | col3 < 100
sum(row2 == 1)
which(row2 == 1)
```


```{r quiz}
quiz(
  question("The sum line does the following:",
    answer("counts how many elements are in row2"),
    answer("counts how many elements are not equal to 1 in row2"),
    answer("counts how many 1s there are in row2", correct = TRUE),
    answer("sums the elements in row2"),allow_retry = TRUE,
    random_answer_order = T
  ),
  question("The which line does the following:",
    answer("sums the elements in row2 that are equal to 1"),
    answer("counts how many elements are equal to 1 in row2"),
    answer("tells the positions of the elements in row2 that are equal to 1", correct = TRUE),
    answer("tells the positions of the elements in row2 that are not equal to 1"),allow_retry = TRUE,
    random_answer_order = T
  )
)
```


We can use these conditions to select subsets.

Return the values in `col1` where the corresponding values of `col3` equal 1.

```{r subset1, exercise = TRUE}
col1 <- c(4, 5, 6, 4)
col3 <- c(1, 1, 3, 2)
col1[___]
```

```{r subset1-solution}
col1 <- c(4, 5, 6, 4)
col3 <- c(1, 1, 3, 2)
col1[col3 == 1]
```

```{r subset1-check}
# check code
gradethis::grade_code()
```


We can analyze these subsets as well. Compute the average of only the `col3` values for whom the corresponding `col1` values are 4.

```{r subset2, exercise = TRUE}
col1 <- c(4, 5, 6, 4)
col3 <- c(1, 1, 3, 2)
mean(___)
```

```{r subset2-solution}
col1 <- c(4, 5, 6, 4)
col3 <- c(1, 1, 3, 2)
mean(col3[col1 == 4])
```

```{r subset2-check}
# check code
gradethis::grade_code()
```



Find the minimum of the values in col1 whose col3 values are less than 15.

```{r subset3, exercise = TRUE}
col1 <- c(4, 5, 6, 4)
col3 <- c(10, 10, 30, 2)
min(___)
```

```{r subset3-solution}
col1 <- c(4, 5, 6, 4)
col3 <- c(10, 10, 30, 2)
min(col1[col3 < 15])
```

```{r subset3-check}
# check code
gradethis::grade_code()
```


*Note:* when your data values are text, i.e. letters or words, to use the == sign or any of these other logical operators, you need to put quotes around the text. For example, `col1 == "A"`.

## Data Frames

This data are the result of a survey completed by an advertising firm interested in determining how much to emphasize television advertising in a certain county.  The county contains two towns, labeled 1 and 2, and a rural area, labeled 3.  Town 1 is built around a factory and most households contain factory workers with school-aged children.  Town 2 is an exclusive suburb of a city in a neighboring county and contains older residents with few children at home.  A wide range of individuals and family types live in the rural area.  Residents were asked to estimate the average number of hours each week that the household watched TV.

```{r loadData, exercise = TRUE, warning = FALSE}
tv_data <- read_table(here("inst/extdata/TV_Viewing.txt")) ## use when data is in .txt form

head(tv_data) ## shows the first 6 rows of the data frame tv_data
```

Suppose we want the number of TV viewing hours for each area surveyed. To do so, we will need to subset the data to include only viewing hours from one category of Town at a time.

We could do this using logicals like above. We will need the dollar sign notation to access columns of a data frame. The term before the dollar sign is the dataset name, and the term after the dollar sign is the column name. 

```{r subsetDataL, exercise = TRUE}
tv_data$View_hrs[hours$Town == 1] ## look at View_hrs column in tv_data dataset for only those where the Town id is 1
```

It may also be convenient to use the `subset` function.

```{r subsetDataS, exercise = TRUE}
subset(tv_data, Town == 1) ## first argument is data set, second argument is the logical used to subset
```


## Submit progress

As you have gone through this tutorial, your progress has been kept track of in something called a "hash". I can use this to make sure you have executed all of the code chunks and answered all of the questions. I will use this to give you a completion grade for this assignment. 

If a Google Form window does not appear, you can access it [here](https://docs.google.com/forms/d/e/1FAIpQLSeW9exymQeRir3-xsUfzm6JPVWB9XHaKHkAa-hpR0ne1g7gyw/viewform).

```{r context="server"}
learnrhash::encoder_logic(strip_output = T)
```

```{r encode, echo=FALSE}
learnrhash::encoder_ui(
  ui_before = div(strong("Submit your hash in the form below."), br(), br()),
  ui_after  = learnrhash::iframe_ui(
    src = "https://docs.google.com/forms/d/e/1FAIpQLSeW9exymQeRir3-xsUfzm6JPVWB9XHaKHkAa-hpR0ne1g7gyw/viewform", 
    width="900px", height= "1000px"
  )
)
```



